---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# microbiomer

<!-- badges: start -->
<!-- badges: end -->

The goal of `microbiomer` is to provide tools to more seamlessly integrate tools/functions from both the `phyloseq` and `tidyverse`-packages.

**Note this is a package in development; although the functions were not yet tested extensively within the context of this package, they were tested across several projects as source code. Nevertheless, use these functions at your own risk and please file a report if you come across issues.**

## Installation

You can install the released version of `microbiomer` from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("microbiomer")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("wsteenhu/microbiomer")
```
## Example

### Load packages

```{r}
library(microbiomer); library(phyloseq)
library(tidyverse)
```

### Load data

```{r load}
data(ps_NP)
ps_NP
```

The package includes nasopharyngeal microbiota data collected at 1 week, 1 month and 1 year of life. For more information, see `?ps_NP`.

### Convert

```{r}
otu_table(ps_NP) %>%
  to_RA() %>%
  .[1:5,1:3]
```

`to_RA` converts the otu_table()-object nested within a phyloseq-objected into a total-sum-scaled table (relative abundances instead of raw reads).

### Extract metadata

```{r}
ps_NP %>%
  meta_to_df() %>%
  head()
```

`meta_to_df()` convers the sample_data()-objects within a phyloseq-object into a tidyverse-formatted data table. Also try `otu_tab_to_df()` and `ps_to_df()`

### Prepare for plotting

```{r prep_plot}
ps_NP %>%
  prep_bar(n = 5) %>%
  head()
```

The function `prep_bar()` converts your phyloseq-object into an object that can be readily used for plotting. It includes all metadata-columns, a column 'OTU' (with formatted OTU-names) and a column 'value' with read counts/relative abundances (depending on the format of the otu_table).

### Barplot

```{r plot, out.width = '70%', width = 7, h = 4.8}
ps_NP %>% 
  subset_samples(., birth_mode == "vag" & visit_name == "w1") %>% # select samples
  to_RA %>%
  create_bar()
```

Note this plotting function can be used in conjunction with other ggplot2-functions/extensions, such as coord_flip()/ggforce::facet_col() and facet_wrap().
